---
title: "PA_template"
output: 
  html_document: 
    keep_md: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r preprocessing, echo=TRUE}

# Unzip the .zip file
library ("downloader")
url = 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
download(url, dest="dataset.zip", mode="wb") 
unzip ("dataset.zip", exdir = "C://Users//ktkni//OneDrive//Desktop//project1")

# Import text file.
data = read.csv("activity.csv")
```
## What is mean total number of steps taken per day?  Calculate the sum, mean, and median of the number of steps.  Create a histogram of the frequency of the sum of the number of steps per day.

```{r meantotal, echo=TRUE}

#Calculate the total number of steps taken per day

library (dplyr)

sumofsteps <- data %>% group_by (date) %>% summarise (sum = sum (steps))

sumofstepssum <- sumofsteps$sum

# Calculate the total number of steps taken per day

sum = sum (sumofstepssum, na.rm = TRUE)
mean = mean (sumofstepssum, na.rm = TRUE)
median = median (sumofstepssum, na.rm = TRUE)

print (paste0("This is the sum:", sum))

print (paste0("This is the mean:", mean))

print (paste0("This is the median:", median))

```


```{r histogram, echo=TRUE}
hist (sumofstepssum, main = "Steps per Day", xlab = "Steps")
```

## What is the average daily activity pattern?  Which 5-minute interval contains the maximum number of steps?

```{r averagenumber, echo=TRUE}

#Calculate the total number of steps taken per day

library (dplyr)

averageofinterval <- data %>% group_by (interval) %>% summarise (average = mean (steps, na.rm = TRUE))

plot (averageofinterval$interval, averageofinterval$average, type = "l", xlab = "5 Minute Intervals", ylab = "Average Steps Per Interval")


```

```{r greatestinterval, echo=TRUE}

#Calculate the total number of steps taken per day

library (dplyr)

averageofinterval <- data %>% group_by (interval) %>% summarise (average = mean (steps, na.rm = TRUE))

maxaverage <- max (averageofinterval$average)

highestinterval <- averageofinterval %>% filter (average == maxaverage)

print (paste0("This is the interval that the highest average steps belong to: ", highestinterval$interval))

```
## Imputing missing values. First, calculate the mean of the NA.  Then, replace the NA with the mean of the interval.

```{r numberofNA, echo = TRUE}
numberofNA <- sum(is.na(data))

print (paste0("This is the number of NA: ", numberofNA))
```

```{r replaceNAfunction, echo = TRUE}
MeanIntervalReplaceNA<-function(interval){
    averageofinterval[averageofinterval$interval==interval,]$average
}
```

```{r replaceNA, echo=TRUE}
datawithoutna <- data

for(i in 1:nrow(datawithoutna)){
    if(is.na(datawithoutna[i,]$steps)){
        datawithoutna[i,]$steps <- MeanIntervalReplaceNA(datawithoutna[i,]$interval)
    }
}
```

```{r calculatesumNA, echo = TRUE}
numberofwithoutNA <- sum(is.na(datawithoutna))

print (paste0("This is the number of NA: ", numberofwithoutNA))
```

```{r averagetotalNoNA, echo=TRUE}

#Calculate the total number of steps taken per day

library (dplyr)

sumofstepswithoutna <- datawithoutna %>% group_by (date) %>% summarise (sum = sum (steps))

# Calculate the total number of steps taken per day

sum1 = sum (sumofstepswithoutna$sum)
mean1 = mean (sumofstepswithoutna$sum)
median1 = median (sumofstepswithoutna$sum)

print (paste0("This is the sum:", sum1))

print (paste0("This is the mean:", mean1))

print (paste0("This is the median:", median1))

hist (sumofstepswithoutna$sum, main = "Steps per Day", xlab = "Steps")
```

## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays/weekends to dataframe, echo = TRUE}
datawithoutna$date <- as.Date(strptime(datawithoutna$date, format="%Y-%m-%d"))
datawithoutna$day <- weekdays(datawithoutna$date)
```

```{r add name of weekdays/weekends to dataframe using for and if loop, echo = TRUE}

for (i in 1:length(datawithoutna$day)){
  if (datawithoutna$day[i] == "Saturday"|datawithoutna$day[i] == "Sunday"){
    datawithoutna$timeofweek[i] <- "Weekend"
  }else{
    datawithoutna$timeofweek[i] <- "Weekday"
  }
}


```

```{r timeserieslatticegraph, echo = TRUE}

library (dplyr)
library (lattice)

datawithoutnaforgraph <- datawithoutna %>% group_by (timeofweek, interval) %>% summarise (average = mean (steps))


xyplot (average ~ interval|timeofweek, type = 'l', data = datawithoutnaforgraph, layout=c(1, 2), ylab = "Average Number of Steps")
```